(Re-)bridging the anatomy ontologies with SSSOM

### 2024-07-01 presentation.tex
Add back removed slides about the previous system.
We are now aiming for a 25-min talk, so we can afford to put back the
previously removed slides that expand about the limitatons of the
previous integration pipeline (but put the slide with the overview
first, not last).

Also move the slide about the ROBOT implementation back from the
supplementary slides to the main presentation.

\hypersetup{pdfpagemode=UseNone,
  pdftitle={},
  pdfauthor={Damien Goutte-Gattat}}
\AtBeginSection{
  \begin{frame}{Outline}{}
    \tableofcontents[currentsection]
  \end{frame}
}

@@

\section{Limits of the previous system}

\begin{frame}
  \frametitle{Overview of the previous integration system}

  \begin{columns}
    \begin{column}{.46\textwidth}
      \begin{tikzpicture}[node distance=.3cm]
        \node [Entity,text width=1.3cm,alt=<1>{fill=GenericHighlight}{}] (uberon) {Uberon\\(incl. xrefs)};
        \node [Process,below right=of uberon,alt=<3>{fill=GenericHighlight}{}] (script) {Custom script};
        \node [Entity,above right=of script,align=left] (ssao) {Taxon-specific\\
                                                                ontologies};
        \node [Entity,below=of script,alt=<2>{fill=GenericHighlight}{}] (cob) {Bridge files};
        \node [Process,below=of cob] (merge) {OWL merge};
        \node [Entity,below=of merge] (cm) {Multi-species ontology};
        \draw [->] (uberon.east) -| (script.north) (script.south) -- (cob.north);
        \draw [->] (cob) -- (merge) (merge) -- (cm);
        \draw [->] (uberon.south) |- (merge.west);
        \draw [->] (ssao.south) |- (merge.east);
      \end{tikzpicture}
    \end{column}
    \begin{column}{.51\textwidth}
      \begin{block}{Mappings maintained as \textcolor<1>{red}{xrefs}}\scriptsize\ttfamily
[Term]

@@

  \begin{itemize}
    \item Mappings not available to third parties
    \item \textcolor<3>{red}{Custom pipeline} highly Uberon-specific, inflexible and not reusable
  \end{itemize}

  \note<1>{Mappings are represented and maintained as \emph{cross-references}
  that are carried by the taxon-neutral terms in Uberon and that point to the
  taxon-specific terms in other ontologies -- as shown here, where Uberon's
  ``dorsal vessel heart'' term carries a ``xref'' to FBbt:00003154, which is
  FBbt's term for the fly heart.

  When building Uberon, those cross-references are extracted by a custom
  script that turns them into \emph{bridge files}, which contains OWL
  materialisations of the mappings, in the form of equivalence axioms (for
  example here, an axiom stating the FBbt's ``heart'' is equivalent to an
  Uberon's ``dorsal vessel heart'' that is part of the \emph{Drosophila
  melanogaster} taxon).

  This system has severe limitations, one of the most important being that the
  mappings are directly embedded within the ontology itself and are not
  independently available. In FAIR terms, this makes the mappings much more
  difficult to find, access and reuse than we would like.

  Furthermore, the custom script that generates the bridge files from the
  cross-references is highly specific to Uberon; its behaviour are completely
  hardcoded to fit Uberon's use case, so the integration logic is difficult to
  modify and almost impossible to reuse by other projects.}

@@

  \begin{enumerate}
    \item Cross-ontology mappings (COM) as first-class entities
      \begin{itemize}
        \item available independently of the ontology itself
      \end{itemize}
    \item Represent mappings using a standard format intended for that purpose
      \begin{itemize}
        \item allowing for mapping-specific metadata
        \item allowing for meaningful mapping predicates
      \end{itemize}
    \item Make it easy to derive OWL bridge files from mappings
      \begin{itemize}
        \item without requiring Uberon-specific tooling
      \end{itemize}
  \end{enumerate}

@@

  First, we want to treat mappings as ``first-class entities'', on the same
  level as the terms of the ontology. This notably means that they must be
  publicly available on their own, without requiring people interested in them
  to forcefully extract them from the ontology.

  Second, we want to store the mappings using a format that is both standard
  and specifically intended for representing mappings. That format should make
  it possible to store mapping-specific metadata, and to give precise meaning
  to a mapping (not just saying that two terms are mapped, but: who asserted
  that those two terms should be mapped? When? On what basis? And what does
  the mapping actually means?).

@@

\section{Implementation}

@@

\begin{frame}
  \frametitle{New mapping predicates for cross-species mappings}

@@

  \note{One of the benefits of the SSSOM standard is that it allows to specify
  an explicit mapping predicate, so as to give a precise meaning to each
  mapping. It is then possible to distinguish, for example between mappings of
  terms that represent exactly the same concept in two different vocabularies,
  and mappings of terms that represent \emph{similar}, but not exactly
  identical, concepts.

@@

@@

  First, a simple rule with a simple filter that selects all mappings where
  the subject is a Uberon term, and apply an inversion preprocessor to them.
  Inversion flips a mapping around the predicate, so that the subject becomes
  the object and the other way round. So after applying this rule, all
  mappings with a Uberon subject in the set will become mappings with a Uberon
  object.

@@

@@

  \begin{columns}
    \begin{column}{.45\textwidth}
      \begin{center}
        \begin{tikzpicture}[node distance=.5cm]
          \node [Entity] (uberon) {Uberon};
          \node [Entity,right=of uberon] (com) {Mapping sets};
          \node [Entity,right=of com,align=left] (ssao) {Taxon-specific\\
                                                         ontologies};
          \node [Process,below=of com,alt=<2>{fill=GenericHighlight}{}] (sssomt) {SSSOM/T-OWL};
          \node [Entity,below=of sssomt] (cob) {Bridge files};
          \node [Process,below=of cob,alt=<3>{fill=GenericHighlight}{}] (merge) {OWL merge};
          \node [Entity,below=of merge] (cm) {Multi-species ontology};
          \draw [->] (com) -- (sssomt) (sssomt) -- (cob);
          \draw [->] (cob) -- (merge) (merge) -- (cm);
          \draw [->] (uberon.south) |- (merge.west);
          \draw [->] (ssao.south) |- (merge.east);
        \end{tikzpicture}
      \end{center}
    \end{column}
    \begin{column}{.45\textwidth}
      \begin{enumerate}
        \item Generating the bridge files through a \textcolor<2>{red}{SSSOM/T-OWL ruleset}\\
              (\texttt{robot sssom:inject})
        \item \textcolor<3>{red}{Merging} Uberon, the bridges, and the taxon-specific ontologies\\
              (\texttt{robot merge})
      \end{enumerate}
    \end{column}
  \end{columns}

  \note<1>{This leads us to our new, SSSOM-based integration pipeline for Uberon.
  That pipeline has three distinct inputs: (1) Uberon itself, (2) the mappings
  between Uberon and the taxon-specific ontologies (which are no longer
  embedded within the ontology itself), and (3) the taxon-specific ontologies.

  The bridge files are generated by a ROBOT plugin that implements the
  SSSOM/T-OWL language, and a custom SSSOM/T-OWL ruleset that contains the
  Uberon-specific logic, in the form of rules similar to the one seen in the
  last example.

  After this step, we have only OWL files, which can all be merged using a
  standard ROBOT merge operation, producing the multi-species ontology that we
  want.

  Overall, that pipeline is both more robust, as most of the code is
  consolidated in ROBOT, and more flexible, as the Uberon-specific logic is
  easily modifiable in the SSSOM/T-OWL ruleset.}

@@

  \note{Beyond the more robust and more flexible pipeline, the new approach
  benefits for downstream users of Uberon as well.

  The mappings between Uberon and the taxon-specific ontologies are now
  first-class artefacts. They can have their own metadata, allowing for
  traceability and accountability of each individual mapping statement. They
  can be maintained separately from the Uberon ontology itself, for example to
  delegate them to the taxon-specific ontologies themselves (something we have
  already started doing for the mappings with FBbt, which are now maintained
  by FBbt rather than by Uberon).  They are available as a distinct release
  product, which is findable on its own, in a format that is still relatively
  new but that we hope will become the standard format for semantic mappings.

  And for everyone else, we now have a generic framework to integrate multiple
  ontologies with SSSOM-maintained mappings. The SSSOM/T-OWL language, and its
  implementation in the SSSOM ROBOT plugin, has no ties whatsoever to Uberon
  and can be reused by any project. Only the SSSOM/T-OWL rules that we use are
  tailored to the needs of Uberon.}

@@

\begin{frame}[noframenumbering]
  \frametitle{Implementation in ROBOT commands}

  \begin{block}{sssom:xref-extract command}
    Convert existing mappings represented as cross-references into a SSSOM set

    \medskip
    {\ttfamily\scriptsize
robot sssom:xref-extract -i input.owl --mapping-file output.sssom.tsv
    }
  \end{block}

  \begin{block}{sssom:inject command}
    Apply SSSOM/T-OWL rules to a mapping set and inject the resulting axioms into an ontology

    \medskip
    {\ttfamily\scriptsize
robot sssom:inject -i input.owl --sssom mappings.sssom.tsv --ruleset bridge.rules
    }
  \end{block}

  \note{To actually implement our new approach, we first, as a preliminary
  step, added plugin support to ROBOT, the main tool we use to manipulate OWL
  and OBO ontologies. This is so that we can quickly add new commands to
  ROBOT. We then developed a SSSOM plugin for ROBOT, which provides two
  distinct commands.

  The first is a command to facilitate the transition to SSSOM-maintained
  mappings by automatically extracting mappings stored as cross-references
  within an ontology into a mapping set in the SSSOM/TSV format.

  More importantly, the second is a command that implements the SSSOM/T-OWL
  domain-specific language. That command takes a mapping set and a set of
  SSSOM/T-OWL rules as input, and will inject into the current ontology the
  axioms produced by applying the SSSOM/T-OWL rules to the mapping set.}

\begin{frame}[fragile,noframenumbering]
  \frametitle{Example: FBbt/Uberon/CL integration test}

  A single ROBOT command to merge FBbt with Uberon, CL, and the mappings-derived
  bridging axioms, and check for possible inconsistencies that would reveal
  incorrect mappings:

  \begin{block}{}
\begin{verbatim}
robot merge -i fbbt.owl -i uberon.owl -i cl.owl \
      sssom:inject --sssom mappings.sssom.tsv   \
                   --ruleset bridging.rules     \
      reason -r ELK
\end{verbatim}
  \end{block}

  \note{Lastly, here is an example to illustrate some of the benefits of the
  approach for third-party users. As I've mentioned, at FlyBase we are now
  maintaining ourselves the mappings between FBbt and Uberon.

  Because FBbt and Uberon are developed separately, there is always a risk
  that, whenever something changes in either FBbt, Uberon, or the
  FBbt-to-Uberon mappings, an inter-ontology inconsistency is introduced.
  Testing to detect and prevent such inconsistencies is difficult and,
  historically, has never actually been done on a regular basis.

  But with the SSSOM approach, such testing can be done with a single ROBOT
  command in three steps, as shown here: (1) we merge the current versions of
  FBbt, Uberon, and CL; (2) we derive OWL axioms from the FBbt-to-Uberon/CL
  mappings and inject them into the merged ontology; (3) we reason over the
  merged ontology to detect inconsistencies. This takes but a couple of
  minutes, and that test is now part of our test suite in FBbt. Thanks to this
  approach, over the past two years we have rooted out all existing
  inconsistencies between FBbt and Uberon.}

@@

>>>>

\hypersetup{pdfpagemode=UseNone,
  pdftitle={},
  pdfauthor={Damien Goutte-Gattat}}
\AtBeginSection{
  \begin{frame}{Outline}{}
    \tableofcontents[currentsection]
  \end{frame}
}
\AtBeginSubsection{
  \begin{frame}{Outline}{}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

@@

\section{Existing system and its limitations}

\begin{frame}
  \frametitle{Integration pipeline overview}

  \begin{columns}
    \begin{column}{.46\textwidth}
      \begin{tikzpicture}[node distance=.3cm]
        \node [Entity,text width=1.3cm,alt=<1>{fill=GenericHighlight}{}] (uberon) {Uberon\\(incl. xrefs)};
        \node [Process,below right=of uberon] (script) {Custom script};
        \node [Entity,above right=of script,align=left] (ssao) {Taxon-specific\\ontologies};
        \node [Entity,below=of script,alt=<2>{fill=GenericHighlight}{}] (cob) {Bridge files};
        \node [Process,below=of cob] (merge) {OWL merge};
        \node [Entity,below=of merge] (cm) {Multi-species ontology};
        \draw [->] (uberon.east) -| (script.north) (script.south) -- (cob.north);
        \draw [->] (cob) -- (merge) (merge) -- (cm);
        \draw [->] (uberon.south) |- (merge.west);
        \draw [->] (ssao.south) |- (merge.east);
      \end{tikzpicture}
    \end{column}
    \begin{column}{.53\textwidth}
      \begin{block}{Mappings maintained as \textcolor<1>{red}{xrefs}}\scriptsize\ttfamily
[Term]

@@

  \note<1>{Here is a brief overview of the integration pipeline.

  Mappings between Uberon and the taxon-specific ontologies are maintained as
  cross-references that are carried by the taxon-neutral terms in Uberon and
  that point to the taxon-specific terms in other ontologies -- as shown here,
  where Uberon's ``dorsal vessel heart'' term carries a ``xref'' to
  FBbt:0003154, which is FBbt's term for the fly heart.

  At build time, xrefs are extracted by a custom script that turns them into
  bridge files, which contains OBO materialisations of the mappings, in the
  form of equivalence axioms (such as an axiom stating that FBbt's ``heart''
  is equivalent to a Uberon ``dorsal vessel heart'' that is part of the
  \emph{Drosophila melanogaster} taxon).

  While the process looks superficially simple (though it is, in fact, quite
  complex, it's just that most of the complexity is hidden behind that
  script), it is also severely limited.

  The limitations stem both from the way the mappings are represented, and the
  way they are transformed into bridge files.}

\begin{frame}
  \frametitle{Issues with mappings maintained as cross-references}

  \begin{block}{Mappings embedded within the ontology}
    \begin{itemize}
      \item Not available separately
     \end{itemize}
  \end{block}

  \begin{block}{No additional data}
    \begin{itemize}
      \item All the cross-references say is that the mappings exist
    \end{itemize}
  \end{block}

  \begin{block}{Meaningless mapping predicate}
    \begin{itemize}
      \item Overloading of the \emph{oboInOwl:hasDbXref} property
      \item Workaround: {\ttfamily\scriptsize treat-xrefs-as-reverse-genus-differentia FBbt NCBITaxon:7227}
        \begin{itemize}
          \item Assigns a meaning to \emph{all} cross-references between FBbt and Uberon
        \end{itemize}
    \end{itemize}
  \end{block}

  \note{The one benefit of using xrefs to represent mappings is that it is a
  very simple method, but that simplicity comes at a high cost.

  First, the mappings are embedded within the ontology and as such are not
  independently available -- they need to be forcefully extracted whenever
  someone needs them. This makes the mappings more difficult to find, access,
  and reuse than it should be.

  Second, the cross-reference is telling us that there is a mapping, but we
  can't know anything else about that mapping. Why does that mapping exist?
  Who asserted it? When? On what basis? This is simply not recorded.

  Third, the \emph{oboInOwl:hasDbXref} annotation has been used for a lot of
  things since its introduction, and no longer has a clear and unambiguous
  meaning -- what it means depends on where it is used. At best, it indicates
  that two terms are mapped \emph{somehow}, but the nature of the relation is
  unspecified.

  In Uberon we can circumvent that last issue by using supplementary
  ontology-level annotations to provide meaning to some cross-references. But
  this is more a workaround than a proper solution, and it comes with problems
  of its own.}

\begin{frame}
  \frametitle{Issues with the bridge generation process}

  \begin{columns}
    \begin{column}{.4\textwidth}
      \begin{center}
        \begin{tikzpicture}
          \node [File] (uberon) at (0,0)    {uberon.obo};
          \node [File] (script) at (0,-1)   {make-bridge-ontology-from-xrefs.pl};
          \node [File] (bridge) at (0,-2)   {uberon-bridge-to-fbbt.obo};
          \node [File]          at (0,-2.5) {uberon-bridge-to-wbbt.obo};
          \node [File]          at (0,-3)   {uberon-bridge-to-ma.obo};
          \node [File]          at (0,-3.5) {uberon-bridge-to-....obo};
          \draw[->] (uberon) -- (script) (script) -- (bridge);
        \end{tikzpicture}
      \end{center}
    \end{column}
    \begin{column}{.55\textwidth}
      \begin{itemize}
        \item Lack of flexibility
        \item Maintenance nightmare
        \item Dependent on the OBO format
        \item Highly Uberon-specific
      \end{itemize}
    \end{column}
  \end{columns}

  \note{The bridge files are generated by this custom Perl script, which reads
  the Uberon ontology file and generates one bridge file for each
  taxon-specific ontology. That script is plagued with several issues.

  The process is very rigid. For example, it is impossible to generate
  only one bridge file on-demand. Almost everything is hardcoded in the
  script, making it difficult to change anything whenever needed.

  The script is difficult to maintain for most Uberon editors and developers,
  in no small part because it conflates the technical aspects of extracting
  the cross-references and generating the bridges, and the logical aspects of
  determining which bridging axioms should be produced.

  It's a smaller issue, but the script also only works with the OBO files,
  both for its input and its output.

  But the main problem is that the bridge generation script is highly specific
  to Uberon. It is almost impossible to reuse its logic elsewhere. Yet the
  need for integrating several ontologies together, using mappings as glue to
  connect terms between those ontologies, is certainly not a Uberon-specific
  need.}

@@

  \begin{enumerate}
    \item Treat cross-ontology mappings as proper entities
      \begin{itemize}
        \item Mappings set made available as independent artefact
      \end{itemize}
    \item Represent COMs using a standard format intended for that purpose
      \begin{itemize}
        \item allowing for mapping-specific metadata
        \item allowing for meaningful mapping predicates
      \end{itemize}
    \item Make it easy to derive bridging axioms from mappings
      \begin{itemize}
        \item without requiring Uberon-specific tooling
      \end{itemize}
  \end{enumerate}

@@

  First, we want the cross-ontology mappings to be treated as first-class
  objects, similarly to the terms of the ontology, rather than merely
  annotations on the terms. They should in fact be treated as the source
  artefact from which the bridge files are derived. As with other source
  artefacts, they must be published on their own and be independently findable.

  Second, we want to store the mappings using a format that is both standard
  and specifically intended for representing mappings. That format should make
  it possible to store mapping-specific metadata, and to give precise meaning
  to every individual mapping (everything that is not possible with the
  cross-reference format).

@@

\section{Implementation of a new approach}

\subsection{New representation of mappings}

@@

\begin{frame}
  \frametitle{Meaningful predicates for cross-species mappings}

@@

  \note{One of the benefits of the SSSOM model is that it allows to specify
  an explicit mapping predicate, so as to give a precise meaning to each
  mapping. It is then possible to distinguish, for example between mappings of
  terms that represent exactly the same concept in two different vocabularies,
  and mappings of terms that represent \emph{similar}, but not exactly
  identical, concepts.

@@

\subsection{New bridge generation process}

@@

  First, a simple rule with a simple filter that selects all mappings where
  the subject is a Uberon term, and applies an inversion preprocessor to them.
  Inversion flips a mapping around the predicate, so that the subject becomes
  the object and the other way round. So after applying this rule, all
  mappings with a Uberon subject in the set will become mappings with a Uberon
  object.

@@

\begin{frame}
  \frametitle{Adding SSSOM/T-OWL support to ROBOT}

  \begin{block}{Prerequisite step}
    Support for ``pluggable commands'' in ROBOT (since version 1.9.5)
  \end{block}

  \begin{block}{sssom:inject command}
    Apply \textcolor<2>{red}{SSSOM/T-OWL} rules to \textcolor<3>{red}{a
    mapping set} and inject the resulting axioms into the
    \textcolor<4>{red}{current ontology}

    \medskip
    {\ttfamily\scriptsize
robot sssom:inject -i \textcolor<4>{red}{input.owl} --sssom \textcolor<3>{red}{mappings.sssom.tsv} --ruleset \textcolor<2>{red}{bridge.rules}
    }
  \end{block}

  \note<1>{With the SSSOM/T-OWL language fully designed, we then needed to
  actually implement it. We chose to add support for SSSOM/T-OWL to ROBOT,
  since this is the main tool we use for most of our pipelines.

  The first step was to add plugin support for ROBOT -- that is, the
  possibility to add third-party commands to ROBOT. This is so that we could
  quickly add new commands, such as SSSOM-related commands, to ROBOT without
  being tied by ROBOT's release cycle. This was done with version 1.9.5 of
  ROBOT, released at the end of Summer 2023.

  We then developed a SSSOM plugin for ROBOT, which among other things,
  provides a command that implement our SSSOM/T-OWL language. That command
  will read a set of SSSOM/T-OWL rules and apply them to a set of mappings,
  thereby producing bridging axioms which will then be injected into the
  ontology that is currently being manipulated by ROBOT.}

@@

  \begin{columns}
    \begin{column}{.45\textwidth}
      \begin{center}
        \begin{tikzpicture}[node distance=.5cm]
          \node [Entity] (uberon) {Uberon};
          \node [Entity,right=of uberon,alt=<2>{fill=GenericHighlight}{}] (com) {Mapping sets};
          \node [Entity,right=of com,align=left,alt=<3>{fill=GenericHighlight}{}] (rules) {SSSOM/T-OWL\\
                                                                                           ruleset};
          \node [Process,below=of com,alt=<4>{fill=GenericHighlight}{}] (inject) {sssom:inject};
          \node [Process,below=of inject,alt=<5>{fill=GenericHighlight}{}] (merge) {merge};
          \node [Entity,right=of merge,align=left] (ssao) {Taxon-specific\\
                                                         ontologies};
          \node [Entity,below=of merge] (cm) {Multi-species ontology};
          \draw [->] (com) -- (inject) (inject) -- (merge) (merge) -- (cm);
          \draw [->] (rules.south) |- (inject.east);
          \draw [->] (uberon.south) |- (inject.west);
          \draw [->] (ssao.west) -- (merge.east);
        \end{tikzpicture}
      \end{center}
    \end{column}
    \begin{column}{.45\textwidth}
      \begin{enumerate}
        \item Maintaining (and collecting) mappings as \textcolor<2>{red}{SSSOM sets}
        \item Maintaining the integration logic as \textcolor<3>{red}{SSSOM/T-OWL rules}
        \item Generating the bridging axioms through \textcolor<4>{red}{SSSOM/T-OWL support in ROBOT}
        \item \textcolor<5>{red}{Merging in} the taxon-specific ontologies
      \end{enumerate}
    \end{column}
  \end{columns}

  \note<1>{This leads us to our new, SSSOM-based integration pipeline for
  Uberon.

  The mappings are now maintained separately from Uberon itself, as SSSOM
  mapping sets.

  The integration logic, which decides how the bridging axioms should be
  generated, is also maintained separately, as a large ruleset written in the
  SSSOM/T-OWL language. It contains rules similar to the ``complex'' example
  of rule shown previously.

  At build time, we use our SSSOM ROBOT plugin to apply that ruleset to our
  mappings, and inject the resulting bridging axioms into Uberon.

  Then, we just need to merge the taxon-specific ontologies into the result of
  the previous step, producing the multi-species ontology that we want.

  Overall, that pipeline is both more robust, as most of the code is
  consolidated in the ROBOT plugin, and more flexible, as the decision logic
  is easily modifiable in the SSSOM/T-OWL ruleset.}

@@

  \note{Beyond the more robust and more flexible pipeline, which is mostly of
  interest for Uberon editors and developers, the new approch has broader
  benefits.

  The mappings between Uberon and the taxon-specific ontologies are now
  first-class artefacts. They can have their own metadata, allowing for
  traceability and accountability of each individual mapping statement, and
  their own mapping predicate. They can be maintained separately from the
  Uberon ontology itself, and we can even delegate the maintenance of some of
  them to the taxon-specific ontologies themselves (something we have already
  started doing for the mappings with FBbt, which are now maintained by
  FlyBase rather than by Uberon).  They are available as a distinct release
  product, which is findable on its own, in a format that is still relatively
  new but that we hope will become the standard format for semantic mappings.

  And beyond Uberon, we now have a generic framework to integrate multiple
  ontologies with SSSOM-maintained mappings. The SSSOM/T-OWL language, and its
  implementation in the SSSOM ROBOT plugin, has no ties whatsoever to Uberon
  and can be reused by any project. Only the SSSOM/T-OWL rules that we use are
  tailored to the needs of Uberon.}

\begin{frame}[fragile]
  \frametitle{Example of reuse: FBbt/Uberon integration test}

  A single ROBOT command to \textcolor<1>{red}{merge FBbt with Uberon},
  \textcolor<2>{red}{inject the mappings-derived bridging axioms}, and
  \textcolor<3>{red}{check for possible inconsistencies} that would reveal
  incorrect mappings:

  \begin{block}{}
\begin{semiverbatim}
robot \textcolor<1>{red}{merge} -i \textcolor<1>{red}{fbbt.owl} -i \textcolor<1>{red}{uberon.owl} \\
      \textcolor<2>{red}{sssom:inject} --sssom \textcolor<2>{red}{mappings.sssom.tsv}   \\
                   --ruleset \textcolor<2>{red}{bridging.rules}     \\
      \textcolor<3>{red}{reason} -r ELK
\end{semiverbatim}
  \end{block}

  \note<1>{Lastly, here is an example to illustrate some of the benefits of
  reusable integration logic.

  Because FBbt and Uberon are developed separately, there is always a risk
  that, whenever something changes in either FBbt or Uberon, an inter-ontology
  inconsistency is introduced.  Because the logic to integrate FBbt into
  Uberon was impossible to reuse outside of Uberon, it was difficult for FBbt
  editors to detect and prevent such inconsistencies whenever we edited the
  ontology.

  But with the SSSOM approach, such testing can be done with a single ROBOT
  command in three steps, as shown here: (1) we merge the current versions of
  FBbt and Uberon; (2) we derive OWL axioms from the FBbt-to-Uberon mappings
  and inject them into the merged ontology; (3) we reason over the merged
  ontology to detect inconsistencies. This takes but a couple of minutes, and
  that check is now part of our test suite in FBbt. Thanks to this approach,
  over the past two years we have rooted out all existing inconsistencies
  between FBbt and Uberon.}

@@

@@

% vim: tw=78 et ai ts=2 st=2 sw=2
